---
title: "CelltypeAnnotation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CelltypeAnnotation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ELeFHAnt)
```
# Celltype Annotation
##### Celltype annotation is a function to annotate celltypes in a single cell datasets

## Requirements
##### It requires a reference dataset (a processed Seurat Object with Celltypes column in metadata) and a query dataset (a processed seurat object with seurat_clusters column in metadata). One can choose from randomForest, SVM or Ensemble classifiction method to learn celltypes from reference dataset and then predict celltypes for query dataset.

## How the function works?
### Merge Reference and Query
##### Merge reference and query into a seurat object
### Normalization
##### Run Log2 Normalization on the merged object using NormalizeData function in Seurat
### Feature Selection
##### Obtain highy variable features from the merged data using FindVariableFeatures function in Seurat
### Scaling
##### Scale the merged object for highly variable features using ScaleData function in Seurat
### Generate train and test data
##### Training set is the reference data with celltypes extracted from scaled data. Test set is the query data with seurat_clusters extracted from scaled data.
### Train Classifiers
##### Training data is used for training randomForest and SVM classifiers.
### Predict Celltypes
##### Each classifier is after training is then used to classify celltypes for the test data
### Ensemble Learning
##### Ensemble learning based classsification uses both randomForest and SVM predictions to define final predictions. It does that by assigning weights (accuracy while learning) to the predictions from each classifier

## How to use the function?
### Load Library ELeFHAnt
##### library(ELeFHAnt)
### Assing parameters in the function
##### CelltypeAnnotation(reference=pbmc, query=mydata, downsample=TRUE, downsample_to = 100, classification.method = "Ensemble", crossvalidationSVM = 10, validatePredictions = TRUE, selectvarfeatures = 2000, ntree = 500, classification.approach = "ClassifyCells")
## What each parameter is?
##### reference: a processed Seurat Object with Celltypes column in metadata
##### query: a processed seurat object with seurat_clusters column in metadata
##### downsample: logical Indicator (TRUE or FALSE) to downsample reference and query enabling fast computation
##### downsample_to: a numerical value > 1 to downsample cells [Default: 100] in reference and query for Celltypes and seurat_clusters resspectively
##### classification.method: choose classification method for learning and predicting celltypes. randomForest (decision trees), SVM (Support Vector Machines) or Ensemble (uses estimation robustness of both randomForest and SVM to predict)
##### crossvalidationSVM: if a integer value k>0 is specified, a k-fold cross validation on the training data is performed to assess the quality of the model
##### validatePredictions: logical indicator (TRUE or FALSE) to asses predictions by deploying gene set enrichment analysis
##### selectvarfeatures: number of variable features to select while training (default: 2000)
##### ntree: number of trees randomForest classifier should build (Default: 500)
##### classification.approach apprach to classify cells 1) ClassifyCells 2) ClassifyCells_usingApproximation. Default: ClassifyCells. We recommend using ClassifyCells_usingApproximation when reference has significantly less number of cells compared to query