---
title: "DeduceRelationship"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DeduceRelationship}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ELeFHAnt)
```
# Deduce Relationship
##### Deduce Relationship is a function to infer relationships among celltypes between two datasets

## Requirements
##### It requires two datasets (processed Seurat Objects with Celltypes column in metadata). One can choose from randomForest, SVM or Ensemble classifiction method to obtain relationships among celltypes. Function outputs a heatmap with celltypes from dataset1 as rows and celltypes from dataset2 as columns..

## How the function works?
### Merge Reference1 and Reference2
##### Merge reference1 and reference2 into a seurat object
### Normalization
##### Run Log2 Normalization on the merged object using NormalizeData function in Seurat
### Feature Selection
##### Obtain highy variable features from the merged data using FindVariableFeatures function in Seurat
### Scaling
##### Scale the merged object for highly variable features using ScaleData function in Seurat
### Generate train and test data
##### Training set is the reference1 and test set is the reference2 extracted from scaled data
### Train Classifiers
##### Training data is used for training randomForest and SVM classifiers.
### Infer Relationship
##### Each classifier is after training is then used to infer relationship among celltypes between reference1 and reference2
### Ensemble Learning
##### Ensemble learning based classsification uses both randomForest and SVM predictions to infer relationship among celltypes between reference1 and reference2. It does that by assigning weights (accuracy while learning) to the predictions from each classifier

## How to use the function?
### Load Library ELeFHAnt
##### library(ELeFHAnt)
### Assing parameters in the function
##### DeduceRelationship(reference1 = NULL, reference2 = NULL, downsample = TRUE, downsample_to = 100, classification.method = c("randomForest", "SVM", "Ensemble"), crossvalidationSVM = 5, selectvarfeatures = 2000)
## What each parameter is?
##### reference1: a processed Seurat Object with Celltypes column in metadata
##### reference2: a processed seurat object with Celltypes column in metadata
##### downsample: logical Indicator (TRUE or FALSE) to downsample reference and query enabling fast computation
##### downsample_to: a numerical value > 1 to downsample cells [Default: 100] in reference and query for Celltypes and seurat_clusters resspectively
##### classification.method: choose classification method for learning and predicting celltypes. randomForest (decision trees), SVM (Support Vector Machines) or Ensemble (uses estimation robustness of both randomForest and SVM to predict)
##### crossvalidationSVM: if a integer value k>0 is specified, a k-fold cross validation on the training data is performed to assess the quality of the model
##### selectvarfeatures: number of variable features to select while training (default: 2000)