---
title: "LabelHarmonization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LabelHarmonization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ELeFHAnt)
```
# Label Harmonization
##### Label Harmonization is a function to harmonize cell labels (celltypes) across single cell datasets.

## Requirements
##### It requires a list of processed Seurat Objects with Celltypes column in metadata or a integrated seurat object (integrated seurat object with Celltypes and seurat_clusters column in metadata). One can choose from randomForest, SVM or Ensemble classifiction method to harmonize celltypes.

## How the function works?
### Integrate seurat objects
##### Integrate list of seurat objects into an integrated object using Seurat's Canonical Correlation Analysis
### Scaling
##### Scale integrated data using ScaleData function in Seurat
### Feature Selection
##### Integration features are set as selected features
### Generate train and test data
##### Training set and test data is extracted using stratification on scaled integrated data. 60% percent of the data is used as training set and rest of the 40 percent as testing set.
### Train Classifiers
##### Training data is used for training randomForest and SVM classifiers.
### Harmonize Celltypes
##### Each classifier after training is then used to classify celltypes for the test data thereby help assign harmonized labels to integration seurat clusters
### Ensemble Learning
##### Ensemble learning based classsification uses both randomForest and SVM predictions to define final predictions. It does that by assigning weights (accuracy while learning) to the predictions from each classifier

## How to use the function?
### Load Library ELeFHAnt
##### library(ELeFHAnt)
### Assing parameters in the function
##### LabelHarmonization(seurat.objects = c(seuratbject1, seuratbject2, seuratbject3, ..), perform_integration = TRUE, downsample = TRUE, downsample_to = 100, classification.method = "Ensemble", crossvalidationSVM = 10, validatePredictions = TRUE, integrated.atlas = NULL, npcs = 30, resolution = 0.5, selectanchorfeatures = 2000)
## What each parameter is?
##### seurat.objects: a list of processed seurat objects with Celltypes column in their respective meta.data to perform integration on.
##### perform_integration: logical Indicator (TRUE or FALSE) to perform integration using list of seurat.objects
##### integrated.atlas: an integrated seurat object with CellTypes and seurat_clusters column in meta.data. Required if: perform_integration = FALSE
##### downsample: logical Indicator (TRUE or FALSE) to downsample seurat objects enabling fast computation
##### downsample_to: a numerical value > 1 to downsample cells [Default: 100]
##### npcs: number of principal components to compute after integration
##### resolution: value of the resolution parameter, decides size of cell communities.
##### classification.method: choose classification method for learning and harmonizing celltypes. Choices: randomForest (decision trees), SVM (Support Vector Machines) or Ensemble (uses learning robustness of both randomForest and SVM to predict)
##### crossvalidationSVM: if a integer value k>0 is specified, a k-fold cross validation on the training data is performed to assess the quality of the model
##### validatePredictions: logical indicator (TRUE or FALSE) to asses predictions by deploying gene set enrichment analysis
##### selectanchorfeatures: number of anchor features to use for integrating datasets (Default: 2000)